Arabidopsis Tandem Deamination Figure Generation

### 2022-06-08 R/ArabidopsisTandemDeaminationFigureGeneration.rmd
Created Arabidopsis tandem deamination notebooks

---
title: "NHF1 Tandem Deamination Figure Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered]
  mismatchType:
    label: "Mismatch Type"
    value: CC>TT
    input: select
    choices: [CC>TT, GG>AA]
---

@@

expansionOffset = 10
cellType = "NHF1"
CPDTimepoints = list(`1h` = NULL, `8h` = NULL, `24h` = NULL, `48h` = NULL)
sixFourTimepoints = list(`5min` = NULL, `20min` = NULL, `1h` = NULL)
minReadLength = 22; maxReadLength = 30

@@

```{r read_simplified_mismatch_data}
# Get file paths for the mismatch by read data
CPDMismatchData = list(rep1 = CPDTimepoints,
                       rep2 = CPDTimepoints)
for (i in seq_along(CPDMismatchData)) {
  for (j in seq_along(CPDMismatchData[[i]])) {
    CPDMismatchData[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
      cellType = cellType, lesion = "CPD", timepoint = names(CPDMismatchData[[i]])[j],
      repitition = names(CPDMismatchData)[i], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}

sixFourMismatchData = list(rep1 = sixFourTimepoints,
                           rep2 = sixFourTimepoints)
for (i in seq_along(sixFourMismatchData)) {
  for (j in seq_along(sixFourMismatchData[[i]])) {
    sixFourMismatchData[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
      cellType = cellType, lesion = "6-4", timepoint = names(sixFourMismatchData[[i]])[j],
      repitition = names(sixFourMismatchData)[i], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

```{r 6-4_mismatch_pos_frequency}
for (i in 1:length(sixFourMismatchData)) {
  for (j in 1:length(sixFourMismatchData[[i]])) {
    plotMismatchPositionFrequencies(sixFourMismatchData[[i]][[j]],
                                    title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                                    names(sixFourMismatchData)[i], 
                                                    names(sixFourMismatchData[[i]])[j], "Mismatch Pos"),
                                                  collapse = ' '))
  }
}
```

```{r read_mismatch_frequency_z-score_tables}
CPDMMFreqZScores = list(rep1 = CPDTimepoints, rep2 = CPDTimepoints)
for (i in seq_along(CPDMMFreqZScores)) {
  for (j in seq_along(CPDMMFreqZScores[[i]])) {
    CPDMMFreqZScores[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(CPDMMFreqZScores)[i],
      timepoint = names(CPDMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}

sixFourMMFreqZScores = list(rep1 = sixFourTimepoints, rep2 = sixFourTimepoints)
for (i in seq_along(sixFourMMFreqZScores)) {
  for (j in seq_along(sixFourMMFreqZScores[[i]])) {
    sixFourMMFreqZScores[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(sixFourMMFreqZScores)[i],
      timepoint = names(sixFourMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

```{r 6-4_mismatch_position_frequencies_by_read_length, fig.width = 9, fig.height = 5}
plotPositionAcrossTimepointAndReadLength(sixFourMismatchData$rep1,
                                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                                         "Mismatch Pos Freq Rep1"),
                                                       collapse = ' '),
                                         zScoreTables = sixFourMMFreqZScores$rep1)
plotPositionAcrossTimepointAndReadLength(sixFourMismatchData$rep2,
                                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                                         "Mismatch Pos Freq Rep2"),
                                                       collapse = ' '),
                                         zScoreTables = sixFourMMFreqZScores$rep2)
plotPositionAcrossTimepointAndReadLength(sixFourMismatchData$rep1, posType = FIVE_PRIME,
                                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                                         "Mismatch Pos Freq Rep1"),
                                                       collapse = ' '),
                                         zScoreTables = sixFourMMFreqZScores$rep1)
plotPositionAcrossTimepointAndReadLength(sixFourMismatchData$rep2, posType = FIVE_PRIME,
                                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                                         "Mismatch Pos Freq Rep2"),
                                                       collapse = ' '),
                                         zScoreTables = sixFourMMFreqZScores$rep2)
```

@@

```{r 6-4_read_length_distributions}
plotReadLengthFrequencies(sixFourMismatchData$rep1,
                          title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                          "Read Length Freq Rep1"), collapse = ' '))
plotReadLengthFrequencies(sixFourMismatchData$rep2,
                          title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                          "Read Length Freq Rep2"), collapse = ' '))
```

```{r read_mismatch_position_statistics}
CPDGroupedStats = list(rep1 = list(three_prime = NULL, five_prime = NULL),
                       rep2 = list(three_prime = NULL, five_prime = NULL))
for (i in seq_along(CPDGroupedStats)) {
  for (j in seq_along(CPDGroupedStats[[i]])) {
    CPDGroupedStats[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(CPDGroupedStats)[i],
      strandPolarity = names(CPDGroupedStats[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}

sixFourGroupedStats = list(rep1 = list(three_prime = NULL, five_prime = NULL),
                           rep2 = list(three_prime = NULL, five_prime = NULL))
for (i in seq_along(sixFourGroupedStats)) {
  for (j in seq_along(sixFourGroupedStats[[i]])) {
    sixFourGroupedStats[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(sixFourGroupedStats)[i],
      strandPolarity = names(sixFourGroupedStats[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

```{r 6-4_position_differences_across_read_lengths, fig.width = 9, fig.height = 5}
plotGroupedPositionStats(sixFourGroupedStats$rep1$three_prime, sixFourGroupedStats$rep1$five_prime, POS_DIFF,
                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                         "Mismatch Pos Diff Rep1"), collapse = ' '),
                         xAxisBreaks = c(minReadLength:maxReadLength))
plotGroupedPositionStats(sixFourGroupedStats$rep2$three_prime, sixFourGroupedStats$rep2$five_prime, POS_DIFF,
                         title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                         "Mismatch Pos Diff Rep2"), collapse = ' '),
                         xAxisBreaks = c(minReadLength:maxReadLength))
```

@@

fivePrimeMismatchCPDNucFreq = list(rep1 = NULL, rep2 = NULL)
for (i in seq_along(fivePrimeMismatchCPDNucFreq)) {
  fivePrimeMismatchCPDNucFreq[[i]] = fread(generateFilePath(
    directory = tandemDeaminationDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMismatchCPDNucFreq)[i],
    strandPolarity = "five_prime", anchored = TRUE, filtering = filtering, includedMismatches = mismatchType
  ))
}

threePrimeMismatchSixFourNucFreq = list(rep1 = NULL, rep2 = NULL)
for (i in seq_along(threePrimeMismatchSixFourNucFreq)) {
  threePrimeMismatchSixFourNucFreq[[i]] = fread(generateFilePath(
    directory = tandemDeaminationDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "6-4", repitition = names(threePrimeMismatchSixFourNucFreq)[i],
    strandPolarity = "three_prime", anchored = TRUE, filtering = filtering, includedMismatches = mismatchType
  ))
}

fivePrimeMismatchSixFourNucFreq = list(rep1 = NULL, rep2 = NULL)
for (i in seq_along(fivePrimeMismatchSixFourNucFreq)) {
  fivePrimeMismatchSixFourNucFreq[[i]] = fread(generateFilePath(
    directory = tandemDeaminationDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "6-4", repitition = names(fivePrimeMismatchSixFourNucFreq)[i],
    strandPolarity = "five_prime", anchored = TRUE, filtering = filtering, includedMismatches = mismatchType
  ))
}
```

@@

```{r 6-4_mismatch_anchored_nucleotide_frequencies, fig.width = 11, fig.height = 9}
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchSixFourNucFreq$rep1, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "3' Anchored Nuc Freq Rep1"), collapse = ' '),
                               secondaryYAxisLabel = "5' Cut Site Distance",
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchSixFourNucFreq$rep2, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "3' Anchored Nuc Freq Rep2"), collapse = ' '),
                               secondaryYAxisLabel = "5' Cut Site Distance",
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)

plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchSixFourNucFreq$rep1, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "5' Anchored Nuc Freq Rep1"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchSixFourNucFreq$rep2, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "5' Anchored Nuc Freq Rep2"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
```

```{r 6-4_mismatch_anchored_purine_frequencies, fig.width = 11, fig.height = 9}
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchSixFourNucFreq$rep1, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "3' Anchored Purine Freq Rep1"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "5' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(threePrimeMismatchSixFourNucFreq$rep2, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "3' Anchored Purine Freq Rep2"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "5' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = THREE_PRIME, showFivePrimeCutSite = TRUE, expansionOffset = expansionOffset)

plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchSixFourNucFreq$rep1, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "5' Anchored Purine Freq Rep1"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
plotNucFreqVsReadLengthBarPlot(fivePrimeMismatchSixFourNucFreq$rep2, 
                               title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                               "5' Anchored Purine Freq Rep2"), collapse = ' '),
                               yAxisLabel = "Purine Frequency", secondaryYAxisLabel = "3' Cut Site Distance",
                               combineNucleotides = c("AG"), combinedNames = c("Purine"),
                               posType = FIVE_PRIME, showThreePrimeCutSite = TRUE, expansionOffset = expansionOffset)
```

@@

fivePrimeMMAnchoredCPDSeqFreq = list(rep1 = CPDTimepoints,
                                     rep2 = CPDTimepoints)
for (i in seq_along(fivePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(fivePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fivePrimeMMAnchoredCPDSeqFreq[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(fivePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG", filtering = filtering, includedMismatches = mismatchType
    ))
  }
}

threePrimeMMAnchoredSixFourSeqFreq = list(rep1 = sixFourTimepoints,
                                          rep2 = sixFourTimepoints)
for (i in seq_along(threePrimeMMAnchoredSixFourSeqFreq)) {
  for (j in seq_along(threePrimeMMAnchoredSixFourSeqFreq[[i]])) {
    threePrimeMMAnchoredSixFourSeqFreq[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(threePrimeMMAnchoredSixFourSeqFreq)[i],
      timepoint = names(threePrimeMMAnchoredSixFourSeqFreq[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "YR", filtering = filtering, includedMismatches = mismatchType
    ))
  }
}

fivePrimeMMAnchoredSixFourSeqFreq = list(rep1 = sixFourTimepoints,
                                         rep2 = sixFourTimepoints)
for (i in seq_along(fivePrimeMMAnchoredSixFourSeqFreq)) {
  for (j in seq_along(fivePrimeMMAnchoredSixFourSeqFreq[[i]])) {
    fivePrimeMMAnchoredSixFourSeqFreq[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "6-4", repitition = names(fivePrimeMMAnchoredSixFourSeqFreq)[i],
      timepoint = names(fivePrimeMMAnchoredSixFourSeqFreq[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG", filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

```{r 6-4_sequence_frequencies, fig.width = 11, fig.height = 9}
plotSequenceFrequencies(threePrimeMMAnchoredSixFourSeqFreq$rep1,
                        title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                        "3' Anchored YR Freq Rep1"), collapse = ' '),
                        secondaryYAxisLabel = "5' Cut Site Distance", showFivePrimeCutSite = TRUE,
                        expansionOffset = expansionOffset, querySequences = c("CG","CA","TG","TA"))
plotSequenceFrequencies(threePrimeMMAnchoredSixFourSeqFreq$rep2,
                        title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                        "3' Anchored YR Freq Rep2"), collapse = ' '),
                        secondaryYAxisLabel = "5' Cut Site Distance", showFivePrimeCutSite = TRUE,
                        expansionOffset = expansionOffset, querySequences = c("CG","CA","TG","TA"))

plotSequenceFrequencies(fivePrimeMMAnchoredSixFourSeqFreq$rep1,
                        title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                        "5' Anchored TGG Freq Rep1"), collapse = ' '),
                        secondaryYAxisLabel = "3' Cut Site Distance", showThreePrimeCutSite = TRUE,
                        expansionOffset = expansionOffset, posType = FIVE_PRIME, querySequences = c("TGG"))
plotSequenceFrequencies(fivePrimeMMAnchoredSixFourSeqFreq$rep2,
                        title = paste(c("6-4", titleCaseFiltering, titleCaseMismatchType,
                                        "5' Anchored TGG Freq Rep2"), collapse = ' '),
                        secondaryYAxisLabel = "3' Cut Site Distance", showThreePrimeCutSite = TRUE,
                        expansionOffset = expansionOffset, posType = FIVE_PRIME, querySequences = c("TGG"))
```

>>>>

---
title: "Arabidopsis Tandem Deamination Figure Generation"
author: "Ben Morledge-Hampton"
date: "5/2/2022"
output: html_document
params:
  bioinformaticsDirectory:
    label: "Bioinformatics Projects Directory"
    value: C:/Bioinformatics_Projects
    input: select
    choices: [C:/Bioinformatics_Projects, F:/Bioinformatics_Projects]
  filtering:
    label: "Filtering"
    value: NONE
    input: select
    choices: [NONE, TGG_filtered]
  mismatchType:
    label: "Mismatch Type"
    value: CC>TT
    input: select
    choices: [CC>TT, GG>AA]
---

@@

expansionOffset = 10
cellType = "Arabidopsis"
CPDTimepoints = list(`ZT2` = NULL, `ZT14` = NULL)
minReadLength = 24; maxReadLength = 30

@@

```{r read_simplified_mismatch_data}
# Get file paths for the mismatch by read data
CPDMismatchData = list(rep1 = CPDTimepoints,
                       rep2 = CPDTimepoints)
for (i in seq_along(CPDMismatchData)) {
  for (j in seq_along(CPDMismatchData[[i]])) {
    CPDMismatchData[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$simplifiedMismatches, ".tsv",
      cellType = cellType, lesion = "CPD", timepoint = names(CPDMismatchData[[i]])[j],
      repitition = names(CPDMismatchData)[i], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

```{r read_mismatch_frequency_z-score_tables}
CPDMMFreqZScores = list(rep1 = CPDTimepoints, rep2 = CPDTimepoints)
for (i in seq_along(CPDMMFreqZScores)) {
  for (j in seq_along(CPDMMFreqZScores[[i]])) {
    CPDMMFreqZScores[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$mismatchFrequencyZScores, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(CPDMMFreqZScores)[i],
      timepoint = names(CPDMMFreqZScores[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

@@

```{r read_mismatch_position_statistics}
CPDGroupedStats = list(rep1 = list(three_prime = NULL, five_prime = NULL),
                       rep2 = list(three_prime = NULL, five_prime = NULL))
for (i in seq_along(CPDGroupedStats)) {
  for (j in seq_along(CPDGroupedStats[[i]])) {
    CPDGroupedStats[[i]][[j]] = fread(generateFilePath(
      tandemDeaminationDataDirectory, dataTypeStrings$mismatchPositionStats, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(CPDGroupedStats)[i],
      strandPolarity = names(CPDGroupedStats[[i]])[j], filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@

@@

fivePrimeMismatchCPDNucFreq = list(rep1 = NULL, rep2 = NULL)
for (i in seq_along(fivePrimeMismatchCPDNucFreq)) {
  fivePrimeMismatchCPDNucFreq[[i]] = fread(generateFilePath(
    directory = tandemDeaminationDataDirectory, dataTypeStrings$nucleotideFrequencyData, ".tsv",
    cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMismatchCPDNucFreq)[i],
    strandPolarity = "five_prime", anchored = TRUE, filtering = filtering, includedMismatches = mismatchType
  ))
}
```

@@

@@

fivePrimeMMAnchoredCPDSeqFreq = list(rep1 = CPDTimepoints,
                                     rep2 = CPDTimepoints)
for (i in seq_along(fivePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(fivePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fivePrimeMMAnchoredCPDSeqFreq[[i]][[j]] = fread(generateFilePath(
      directory = tandemDeaminationDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(fivePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(fivePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "five_prime",
      anchored = TRUE, sequence = "TGG", filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
```

@@
